---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library("data.table")
library(ggplot2)
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = 'repdata_activity.zip', method = "curl")
unzip("repdata_activity.zip",exdir = "data")
```
## Reading csv Data into Data.Table. 
```{r}
activityD <- data.table::fread(input = "data/activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
Total_Steps <- activityD[, c(lapply(.SD,sum, na.rm = TRUE)), .SDcols = c("steps"), by = .(date)] 

head(Total_Steps, 10)
```
## Mea median 
```{r}
Total_Steps[, .(Mean_Steps = mean(steps), Median_Steps = median(steps))]
```


## What is the average daily activity pattern?

```{r}
IntervalD <- activityD[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval)] 

ggplot(IntervalD, aes(x = interval , y = steps)) +
    geom_line(color="blue", size=1) +
    labs(title = "Average Daily Steps", x = "Interval", y = "Average Steps perday")

```


## 5 minutes Interval containing maximum number of steps 
```{r}
IntervalD[steps == max(steps), .(max_interval = interval)]
##print(IntervalD)
```


## Imputing missing values
```{r}
activityD[is.na(steps), .N ]
```
# OR
```{r}
nrow(activityD[is.na(steps),])
```
## Filling in missing values by mean of dataset. 
```{r}
activityD[is.na(steps), "steps"] <- round(activityD[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps")])

```

##Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
data.table::fwrite(x = activityD, file = "data/tidyData.csv", quote = FALSE)
```
# total number of steps taken per day
```{r}
Total_Steps <- activityD[, c(lapply(.SD, sum, na.rm = TRUE)), .SDcols = c("steps"), by = .(date)] 
```

# mean and median total number of steps taken per day
```{r}
Total_Steps[, .(Mean_Steps = mean(steps), Median_Steps = median(steps))]

library(ggplot2)
ggplot(Total_Steps, aes(x = steps)) +
    geom_histogram(fill = "blue", binwidth = 1000) +
    labs(title = "Daily Steps", x = "Steps", y = "Frequency")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
activityD[, dateTime := as.POSIXct(date, format = "%Y-%m-%d")]
activityD[, `Day of Week`:= weekdays(x = dateTime)]

activityD[grepl(pattern = "Monday|Tuesday|Wednesday|Thursday|Friday", x = `Day of Week`), "weekday or weekend"] <- "weekday"
activityD[grepl(pattern = "Saturday|Sunday", x = `Day of Week`), "weekday or weekend"] <- "weekend"
activityD[, `weekday or weekend` := as.factor(`weekday or weekend`)]
```

